Variants on the Flight and Maintenance Planning problem

### 2019-04-16 latex/notes/bounds/bounds.tex
reformatted the bounds document.

The idea is to define ways with which we can understand the problem, based on the initial input data.

@@

        \begin{tabular}{p{30mm}p{125mm}}
            $t \in \mathcal{T}^s_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}a). \\
            $t' \in \mathcal{T}^m_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + M + E^{min}-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
            $t' \in \mathcal{T}^M_t$ &  time periods $t' \leq |\mathcal{T}| - E^{max} - M$ such that $t' \in \{ t + M + E^{min}-1 , ...,  t + M + E^{max}-1 \}$ (figure \ref{fig:gantt_windows}a). \\
            $t \in \mathcal{T}^{m_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}}\}$ (figure \ref{fig:gantt_windows}b). \\
            $t \in \mathcal{T}^{M_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}} , ...,  Rct^{Init}_i \}$ (figure \ref{fig:gantt_windows}b). \\
            $t' \in \mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}c). \\
            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}$ & pairs of time periods $t_1, t_2 \in \mathcal{TJ}_j$ such that $t_2 \ge t_1 + MT_j-1$ and $t \in \{t_1, ..., t_2\}$. \\
            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}$ & pairs of time periods $t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}$ such that $t_1 \in \mathcal{T}^s_{t} \lor t_2 \in \mathcal{T}^s_{t}$. \\
            $(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}$ & triplet composed of mission $j \in \mathcal{JI}_i$ and $t, t' \in \mathcal{TJ}_j$ such that $t' \ge t + MT_j-1$ and $t \ge t_1 + M$ and $t' < t_2$ \\
        \end{tabular}

% \section{Number and possible dates for maintenances}
% \label{number-and-possible-dates-for-maintenances}

% The aircraft initial state already conditions the first maintenance of each aircraft.

% Moreover, using the maximum and minimum distances between maintenances (min and max size for maintenances cycles respectively), \textbf{we can precalculate the possible total number of maintenances that an aircraft will have in the horizon}. This number will probably be one of the following: {[}1{]}, {[}1, 2{]}, {[}2{]}, {[}1, 2, 3{]}. This will depend on the size of the horizon and maintenance conditions.

\section{Range of number of maintenances per period and cluster}
\label{range-of-number-of-maintenances-per-period-and-cluster}

First, for each period $t$, and using the initial states of aircraft, we can calculate the minimum and maximum number of maintenances that an aircraft could have already done (\emph{ended}) before the start of the period. We can then aggregate this for each aircraft type to obtain a lower and an upper bound on the number of maintenances that can be finished before period $t$.

\begin{align}

	& t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \forall i \in \mathcal{I}, |\mathcal{T}|

\end{align}

Secondly, we can calculate the number of maintenances we can fit until period $t$, by looking at missions requirements in terms of number of aircraft. This would be done by aircraft type. We first take the sum of all the non-mission periods over all aircraft $i \in IK_k \forall k \in K$\$ and all periods $t' < t$. Then, we divide this over the duration of the maintenances ($M$). This would an upper bound.

 Thirdly, we can sum all flight-hour needs of missions of type $k$ before time $t$. If we substract this from the initial remaining flight time of the group of aircraft, we are left with a number of hours that need to be obtained via maintenances. If we divide this number over the $H^{max}$ flight hours a maintenance gives, we get a lower bound on the number of maintenances we need to do for those aircraft of type $k$.  So, in this way, we have two lower bounds and two upper bounds per period and aircraft type. We could just get the maximum and the minimum respectively to get some bounds on the number of maintenances until period $t$.  {%

\section{Mission assignment}\label{mission-assignment}}  {%

\section{Mission assignment for initial periods}\label{mission-assignment-for-initial-periods}}

Specially for the initial periods, and based on the initial state, missions' hourly requirements and minimum assignment duration, it is possible to \textbf{precalculate which aircraft will not be able to start a mission in some periods}. We can do this only for the initial periods because we do not have the actual consumption for individual aircraft.  In the case of having a minimum consumption, this could be used to have some lower bounds too. 
\end{document}

>>>>

The purpose of this document is to present particular bounds that can be derived from the input data for a particular instance of the problem. These derived bounds can be used to prove the infeasibility of a given instance or to add cuts to the original mathematical formulation.

The first section is a description of the input data. The following sections are reserved for specific derived bounds.

@@

        \begin{tabular}{p{30mm}p{125mm}}
            $t' \in \mathcal{T}^M_t$ &  time periods $t' \leq |\mathcal{T}| - E^{max} - M$ such that $t' \in \{ t + M + E^{min}-1 , ...,  t + M + E^{max}-1 \}$. \\
            $t \in \mathcal{T}^{M_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}} , ...,  Rct^{Init}_i \}$. \\
            $(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}$ & triplet composed of mission $j \in \mathcal{JI}_i$ and $t, t' \in \mathcal{TJ}_j$ such that $t' \ge t + MT_j-1$ and $t \ge t_1 + M$ and $t' < t_2$ \\
        \end{tabular}

\section{Bounds on accumulated checks per aircraft and period} \label{sec:rangechecksaircraft}

  For each period $t$, and using the aircraft initial states, we calculate the minimum and maximum number of checks that an aircraft could have already started / ended at the start of the period.

  $TM1^{min}_i$ and $TM1^{max}_i$ represent, respectively, the minimum and maximum periods for starting the first check for aircraft $i$. Analogously, $TM2^{min}_i$ and $TM2^{max}_i$ represent the minimum and maximum period for starting the second check for aircraft $i$.

  \paragraph{Note on the number of checks.} Here we are assuming a maximum number of checks of two. Based on the size of the horizon we are studying (90 - 120 periods) and the minimum - maximum distance between checks (30 to 60 periods), these are the only possibilities.

  These limits are obtained in the following way:

  \begin{align}
  	& TM1^{min}_i = \min_t{\{t \in \mathcal{T}^{M_{Init}}_i\}}   \notag\\
    & TM1^{max}_i = \max_t{\{t \in \mathcal{T}^{M_{Init}}_i\}}   \notag \\
    & TM2^{min}_i = \min_t{\{t \in \mathcal{T}^M_{TM1^{min}_i}\}}\notag \\
    & TM2^{max}_i = \max_t{\{t \in \mathcal{T}^M_{TM1^{max}_i}\}}\notag %\\
  \end{align}

  Lower and upper bounds for the accumulate number of checks started ($M^{Acc}S^{min}_{it}$, $M^{Acc}S^{max}_{it}$) and finished ($M^{Acc}F^{min}_{it}$, $M^{Acc}F^{max}_{it}$) can be pre-calculated based on where the period is located in the planning horizon relative to the possible first and second check ranges and the check duration. We will name the range $M^{Acc}S^{b}_{it}$ to represent both the lower level and upper level of the range of accumulated starts by replacing the terms $min$ and $max$ by $b$, as in $bound$.

  \begin{multicols}{2}
  \[
   M^{Acc}S^{min}_{it} = 
    \begin{cases} 
     0 & \text{if } t \le TM1^{max}_i\\
     1 & \text{if } TM1^{max}_i < t \le TM2^{max}_i \\
     2 & \text{if } t > TM2^{max}_i \\
    \end{cases}
  \]

  \[
   M^{Acc}S^{max}_{it} = 
    \begin{cases} 
     0 & \text{if } t < TM1^{min}_i \\
     1 & TM1^{min}_i \le t < TM2^{min}_i \\
     2 & t \ge TM2^{min}_i \\
    \end{cases}
  \]

  \columnbreak

  \[
   M^{Acc}F^{min}_{it} = 
    \begin{cases} 
     0 & \text{if } t \le TM1^{max}_i + M\\
     1 & \text{if } TM1^{max}_i < t - M \le TM2^{max}_i \\
     2 & \text{if } t > TM2^{max}_i + M \\
    \end{cases}
  \]

  \[
   M^{Acc}F^{max}_{it} = 
    \begin{cases} 
     0 & \text{if } t < TM1^{min}_i + M\\
     1 & TM1^{min}_i \le t - M < TM2^{min}_i\\
     2 & t \ge TM2^{min}_i + M\\
    \end{cases}
  \]

  \end{multicols}

\section{Bounds on accumulated checks at the end of the horizon per aircraft}

  If we evaluate the range of checks per aircraft at the last period ($\mathcal{T}$), we can get one of three options: [1, 1], [1, 2], [2, 2]. Lets call $I^{1M}$ to aircraft that fall into the [1, 1] case and $I^{2M}$ to groups aircraft that fall into the [2, 2] case. These two sets of aircraft can have particular cuts applied to them to guarantee the number of checks.

   If we find ourselves in a case where there is certainty on the number of checks, it is possible to force the variables to reflect this certainty with the following cuts.

  \begin{align}
    & I^{1M} = \{i \in \mathcal{I} \mid M^{Acc}S^{min}_{i|\mathcal{T}|} = M^{Acc}S^{max}_{i|\mathcal{T}|} = 1\} \notag \\
    & I^{2M} = \{i \in \mathcal{I} \mid M^{Acc}S^{min}_{i|\mathcal{T}|} = M^{Acc}S^{max}_{i|\mathcal{T}|} = 2\} \notag
  \end{align}

  \begin{align}
    & m_{it_1|\mathcal{T}|} =  1
      & i \in I^{1M}, t_1 \in \mathcal{T}^{M_{Init}}_i \\
    & m_{it_1|\mathcal{T}|} =  0
      & i \in I^{2M}, t_1 \in \mathcal{T}^{M_{Init}}_i
  \end{align}

  % \[
  % [M^{Acc}S^{min}_{i|\mathcal{T}|}, M^{Acc}S^{max}_{i|\mathcal{T}|}] :
  %   \begin{cases} 
  %    m_{it_1|\mathcal{T}|} =  1
  %     \,\, \forall t_1 \in \mathcal{T}^{M_{Init}}_i, i \in \mathcal{I} & \text{if } [1, 1] \\
  %    m_{it_1|\mathcal{T}|} =  0 \,\, \forall t_1 \in \mathcal{T}^{M_{Init}}_i, i \in \mathcal{I} & \text{if } [2, 2] 
  %   \end{cases}
  % \]

\section{Bounds on mission assignments at the start of the horizon per aircraft} \label{mission-assignment-for-initial-periods}

  Specially for the initial periods, and based on the initial state, missions' hourly requirements and minimum assignment duration, it is possible to \textbf{precalculate which aircraft will not be able to start a mission in some periods}. We can do this only for the initial periods because we do not have the actual consumption for individual aircraft.  In the case of having a minimum consumption, this could be used to have some lower bounds too.

  For each aircraft, we calculate the last period before being able to finish the first possible check ($t^{s}_i$). Before this period, the sum of all assigned misions and consummed flight hours need to be less than the initial remaining flight hours. This is a particular case of the already existing constraints controlling flight hours when assigning checks.

  As with the previous constraints, it is convenient to pre-calculate certain parameters:

  \begin{align}
      & U^{min}_{t_1t_2} = U^{min} \sum_{t=t_1}^{t_2} (t_2 - t) \notag \\
      & H'_{jt_1t_2} = (H_j - U^{min}) (t_2 - t_1) \notag \\
      & H''_{jt_1t_2t'} = \sum_{t=t_1}^{t_2} (H_j- U^{min}) (t' - t) \notag \\
      & t^{s}_i = TM1^{min}_i + M - 1&  \notag
  \end{align}

  The cuts in (\ref{eq:initialFlightHourConsumption}) control the consummed hours before any maintenance can be done on aircraft $i$ (before the period $t^{s}_i$).

  \begin{align}
    & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t^{s}_i}}} a_{ijtt'} H'_{jtt'} + U^{min}_{1t^{s}_i} \leq Rft^{Init}_i
                    & i \in \mathcal{I} \label{eq:initialFlightHourConsumption}
  \end{align}

  Some mission assignments can already be discarded because the equivalent flight hour consumption is more than the initial remaining time:

  \begin{align}
    & a_{ijtt'} = 0 & i \in \mathcal{I}, (j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{i1t^{s}_i} \land Rft^{Init}_i < H'_{jt_1t_2}
  \end{align}

\section{Bounds on accumulated checks per aircraft type and period} \label{sec:rangecheckstype}

  Both $M^{Acc}F^{b}_{it}$ and $M^{Acc}S^{b}_{it}$ previously calculated in section \ref{sec:rangechecksaircraft} can be aggregated by type of fleet $y$.

  \begin{align}
    & YM^{Acc}S^{b}_{yt} = \sum_{i \in \mathcal{IY}_y} M^{Acc}S^{b}_{it} 
        & t \in \mathcal{T}, y \in \mathcal{Y}, b \in \{min, max\} \notag \\
    & YM^{Acc}F^{b}_{yt} = \sum_{i \in \mathcal{IY}_y} M^{Acc}F^{b}_{it} 
        & t \in \mathcal{T}, y \in \mathcal{Y}, b \in \{min, max\} \notag
  \end{align}

  By using the mission required assignments over time, we can also calculate the number of checks we can fit until period $t$: aircraft that are in a mission cannot be in maintenance. Since each mission demands a specific type of aircraft, this bound can be made at the aircraf type level. $JR^{Acc}_{jt}$ represents the accumulated required number of assignments of aircraft for mission $j$ until the end of period $t$.
  We use this value to obtain the sum of all the non-mission periods over all aircraft of type $y$ and all periods $t' \le t$ ($IR^{Acc}_{yt}$). Then, we divide this over the duration of the checks ($M$) to obtain $YM^{Acc}F^{\prime max}_{yt}$, which is an upper bound on the number of checks that can be finished until period $t$ for all aircraft of type $y$.

  \begin{align}
    & JR^{Acc}_{jt} = |\{t' \in \mathcal{TJ}_j \mid t' \le t\}| 
        & t \in \mathcal{T}, y \in \mathcal{Y}\notag\\
    & IR^{Acc}_{yt} = |IY_y| \times t -  \sum_{j \in \mathcal{JY}_y} R_j \times JR^{Acc}_{jt} 
        & t \in \mathcal{T}, y \in \mathcal{Y}\notag\\
    & YM^{Acc}F^{\prime max}_{yt} = \lfloor \frac{YFN_{yt}}{M} \rfloor 
        & t \in \mathcal{T}, y \in \mathcal{Y}\notag
  \end{align}

  By using the mission required flight hours over time, we can calculate how many checks we need until period $t$. We can sum all flight-hour needs of missions of type $y$ until the end of period $t$ ($YH^{Acc}_{yt}$). This demand of flight hours can be subtracted this from the initial remaining flight time of the group of aircraft and then divided over the $H^{max}$ flight hours a check gives. Thus, we get a lower bound $YM^{Acc}F^{\prime min}_{yt}$ on the number of checks we need to do for those aircraft of type $y$ until period $t$.

  \begin{align}
    & YH^{Acc}_{yt} =  \sum_{j \in \mathcal{JY}_y} H_j \times JR^{Acc}_{jt} 
        & t \in \mathcal{T}, y \in \mathcal{Y} \notag\\
    & YM^{Acc}F^{\prime min}_{yt} =  \lceil \frac{YH^{Acc}_{yt} - \sum_{i \in IY_y} Rft^{Init}_i }{H^{max}} \rceil
        & t \in \mathcal{T}, y \in \mathcal{Y} \notag
  \end{align}

  So, in this way, we arrive to two lower bounds and two upper bounds per period and aircraft type. We then get the maximum and the minimum respectively to get some bounds on the number of checks until period $t$. $YM^{Acc}F^{\prime\prime b}_{yt}$ represents the bound $b$ in the number of total checks since the beginning of the planning horizon for all aircraft of type $y$ until period $t$.

  \begin{align}
    & YM^{Acc}F^{\prime\prime min}_{yt} = \max\{YM^{Acc}F^{min}_{yt}, YM^{Acc}F^{\prime min}_{yt}\}\notag\\ 
    & YM^{Acc}F^{\prime\prime max}_{yt} = \min\{YM^{Acc}F^{max}_{yt}, YM^{Acc}F^{\prime max}_{yt}\}\notag
  \end{align}

  \[
    QM^{num}_{tt'} = 
    \begin{cases} 
     1 & t' > t - M \\
     2 & t' \le t - M \\
    \end{cases} 
  \]

  $QM^{num}_{tt'}$ represents the possibility of counting two checks instead of one if both checks start before the period $t$. $t'$ represents the second one of those checks.

  This can be applied via a cut by the following way:

  \begin{align}
    & YM^{Acc}F^{\prime\prime min}_{yt} \le \sum_{\substack{i \in \mathcal{IY}_y, \\ t_1 \in \{1, \ldots, t - M\}, \\ t_2 \in \mathcal{T}^M_{t_1}}} m_{it_1t_2} \times QM^{num}_{tt_2} \le YM^{Acc}F^{\prime\prime max}_{yt}
        & t \in \mathcal{T}, y \in \mathcal{Y} \label{eq:range_maints}
  \end{align}

  Cuts (\ref{eq:range_maints}) limit the starts of checks of aircraft of type $y$ in order to have the number of finished checks to fall between the $YM^{Acc}F^{\prime\prime b}_{yt}$ bounds.

\section{Bounds on accumulated checks per period}

  The bounds on accumulated checks by type and period defined in \ref{sec:rangecheckstype} can be aggregated into the whole fleet.

  \begin{align}
    & M^{Acc}S^{b}_{t} = \sum_{y \in \mathcal{Y}} YM^{Acc}S^{\prime\prime b}_{yt}
        & t \in \mathcal{T}, b \in \{min, max\}  \notag\\
    & M^{Acc}F^{b}_{t} = \sum_{y \in \mathcal{Y}} YM^{Acc}F^{\prime\prime b}_{yt}
        & t \in \mathcal{T}, b \in \{min, max\}  \notag
  \end{align}

  Additionally, the maintenance capacity together with the maintenance duration offer an upper bound on the maximum number of checks that can be finished until a given period $t$.

  \begin{align}
    & M^{Acc}F^{\prime max}_{t} = \lfloor \frac{t}{M} \rfloor \times C^{max}
        & t \in \mathcal{T} \notag
  \end{align}

\end{document}
